\myTitle

### 2024-12-03 cover-letter/responses-to-editor-and-referees.Rmd

---
title: "Responses to Editor and Referees"
output: pdf_document

---

## Editor Comments to Author:

Two reviewers have provided comments on this manuscript and both are favorably impressed.  They have raised a few minor issues that should be addressed in a revision.  In particular, it would be good to better address the limitations when there is inbreeding present.  I look forward to reading a revised version.

_Thank you.  We address the issue of inbreeding in the revised version.  See below.
Our replies and comments are in italic text throughout._

Subject Editor, Molecular Ecology Resources

## Reviewer Comments to Author:

## Reviewer: 1

### General comments:

In this work, Anderson and colleagues present a new R package; gscramble which can be used to simulate genotypes across hybrid classes. The package differs from other commonly used programs such as ADMIXTURE by, sampling alleles without replacement and using recombination rates to account for linkage among large genetic datasets.

Both these considerations are important and represent improvements to current methods. I do have some questions about how the package is implemented and what some of the requirements mean for real-world applications detailed below. Once clarified, I believe this work is suitable for publication in Molecular Ecology Resources.

### Detailed comments:

L90-91: Are the first two figures supposed to appear one after the other in the Introduction? They aren’t referred to in the text until line 202 in the methods.

_Thank you for catching that.  They should not appear there.  That was a byproduct of using the same
\LaTeX{} input for both a typeset-2-column version and our submission version.  We have moved the
figure callouts to their appropriate locations in the paper._

L120-123: Is it not possible to simply filter the loci set based on linkage disequilibrium? Do you have evidence that this will not alleviate the problem?

_No, it is not possible. Genomic material is inherited in chunks that are typically orders of magnitude longer than
the typical range over which LD decays.  Thus, pairs of markers that are
on the same chromosome, are much more likely to be co-inherited than markers that
are not on the same chromosome, even if the two markers are far enough apart on the
same chromosome that they are not in LD. We have noted this in the discussion for any other readers that may
wonder whether LD pruning will solve the problem of physical linkage._

L196-198: How would the requirement of having no inbreeding loops in the GSP affect the applicability of these simulations to real-life scenarios? Within most wild populations, especially threatened species, inbreeding is a regular occurrence.

_This is a reasonable question and we address it now in the Discussion. In short, gscramble works fine
for inbred populations.  Partly this is because if the founders are related (as they would be in a tiny population)
the offspring produced by the GSP will by inbred despite the absence of inbreeding loops in the GSP._

L202-204: Figure captions should not contain required information for the text only. The terminology should be described in the text at the first mention.

_We pulled that long caption out of the figure and include it in the text. That improves things.
Thank you!_

L227-229: Will recombination rates be something the user will have to input?

_Yes, the user has to put those in, though we already note in the discussion that even if an
accurate recombination map is not known, it is still possible to use one simply with
average rates across each chromosome._

L363-365: My understanding was that gscramble accounted for the bias associated with linked loci without the need to filter, so why would an empirical dataset need to be filtered for linked loci? What benefit then does gscramble present over other methods in this regard?

_As noted above, LD and physical linkage are quite different.  Where we now address that in the discussion, we also
note that for applications of the data after it has been run through gscramble, like using ADMIXTURE, it can be
advantageous to have LD-pruned data._

Figure 7: How did the results of gscramble differ from the original paper? How do you know what is the truth? Or do the results simply differ?

_The point of this simulation was to see whether the original authors were justified in limiting their data set to just the 30
markers on separate chromosome arms. We now emphasize in the results section that this choice is justified by the simulation results._

L482: Is there a function in gscramble for users to estimate the “specific-specific recombination rates” from their empirical data?

_No, there is not._

## Reviewer: 2

### Comments to the Author

The manuscript "gscramble: Simulation of admixed individuals without reuse of genetic material" presents a new R package to simulate genetically admixed individuals. The main advancement presented is that this simulation avoids the issue of resampling and the spurious power inflation that can result from this.

The method allows users to upload their own recombination maps and specify complex hybridization scenarios. To test their method the authors present simulations with trout and pig data which showcase the strength of their method.

Overall the paper is well written. It takes complex concepts and presents them in plain english. I commend the authors for the clear organization and concise writing.

The authors do present some of the limitations of their method, but I think they could do a little bit more of this. This might be outside the scope of the paper but I wonder how this might apply in the field of human genetics.

_This is an interesting question. We do feel that gscramble could be applicable in some
contexts within human genetics.  The ready availability of phased dense data in humans
would be appropriate for use with the `preserve_haplotypes = TRUE` option. So, gscramble
could be used to assess how well admixture from different poulations could be detected.
Given the emphasis on molecular ecology in the outlet, however, we agree with the
reviewer that including discussion of human genetics might be outside of the scope of the
paper._

The authors mention "genetic clustering algorithms" multiple times in the text, it seems to be a major motivator for their software, however they only use ADMIXTURE. I wonder if the authors could make these results more robust by presenting results using another clustering algorithm (could even just be K-means) to show that their results hold up using more than other clustering algorithm.

_I think it is fairly well understood from the explanations in Anderson, Waples, and Kalinowski (2008) that
RISPI would also affect something like K-means clustering.  We considered adding more simulations to the paper, but felt that
the paper was already quite long as it is.  We focus on ADMIXTURE in the paper because it provides
an explicit estimate of the admixture fraction, and one of the main points of the paper is that
the variance of admixture fractions within recent hybrids is strongly affected by physical linkage._

The graphics of the paper could be improved with larger text sizes. This is an aesthetic suggestion, but I wonder if the 4 by 4 grid plots could be reduced to a 3 by 3 grid to increase the overall size of each plot while preserving the visualization of the trends.

_Yes, being able to make those 3 x 3 grids could make them more readable; however we do
feel it is important to have those four magnitudes of marker numbers available, since they
do attempt to span the range of the sorts of data sets that might be encountered in molecular
ecology today: 100 = Fluidigm chips and amplicon sequencing; 1000 = marker capture;
10,000 = RAD seq and small SNP chips; 100,000+ RAD seq, large SNP chips and whole genome
sequencing. Ultimately, we chose to keep them 4 x 4 to preserve that range, and symmetry
in the vertical dimension._

## Production Checklist

To hasten the appearance of your manuscript online, I urge you to complete the following prior to returning your revised manuscript:

1. Ensure data is archived in a publicly-accessible repository. Please see the Data Storage and Documentation section in the author guidelines for details of what data should be archived, appropriate public archives for specific data types and suitable file formats. Exemptions to the data archiving policy may be granted for cases involving sensitive information such as human subject data or the location of endangered species.

_Done. Data and code are archived and publicly accessible._

2. Complete Data Accessibility and Benefit-Sharing section in your manuscript. For correct formatting of this section, please see our author guidelines. Please include DOIs or accession numbers in your Data Accessibility Statement. Data Accessibility statements will be checked post acceptance but prior to publication for completeness and that information such as DOIs or accession numbers are included.

_Done. Data and code are archived and publicly accessible, and that is described in the statement._

Molecular Ecology Resources strongly encourages authors to include a Benefit-Sharing Statement, which should disclose non-monetary benefits generated by your research in the context of the Nagoya Protocol (https://www.cbd.int/abs/). Please review the benefits associated with this research and describe them within the Data Accessibility and Benefit-Sharing section in your manuscript. For more information, please see our editorial https://onlinelibrary.wiley.com/doi/10.1111/mec.15702.

_We have added a Benefit-Sharing Statement_

3. Consolidate Supplemental Information into as few files as possible.

    * Use the attached Molecular Ecology Resources Supplemental Information Template if possible (not required)
    * If the file is quite long, create a Table of Contents for the first page
    * Make sure the title of the paper and the names of all authors appear on the Supplemental document.
    * Your Supplemental Information will not be copy-edited.  Do not leave in track-changes and other editing marks.  The document will be posted "as is.”
    * Save as PDF if possible.

_The Supplemental Information has been saved as a PDF_

4. Provide your manuscript in an editable format (not PDF!)

5. Provide tables in an editable format (not images)

6. Make sure all table and figure captions are present and editable

_All will be submitted in LaTeX format, and we will ensure proper rendering_

The title and abstract of your article are very important to increasing its visibility. We encourage you to make your article more discoverable online with the following tips:

https://authorservices.wiley.com/author-resources/Journal-Authors/Prepare/writing-for-seo.html

_Thank you for those SEO tips!_

To submit your revised manuscript:

1. Log in to your Scholar One account and click on the 'manuscripts with decisions' link
2. Click 'create a revision' and follow the on-screen instructions. (Do not upload your revised manuscripts as a new submission.)
3. Please include a point-by-point response to the reviewer comments, including a detailed list of changes made and a rebuttal to any comments with which you disagree, being as specific as possible in your responses.
4. Please provide a version of your manuscript with the changes tracked or highlighted.
5. Your original files will be available to you when you upload your revised manuscript. Please delete any outdated files before completing the submission.

In order to schedule publication efficiently, we impose a deadline of 30 days for the return of revised manuscripts. Manuscripts returned after the deadline may be regarded as new submissions, with a new receipt date, and may be subject to the full peer review process.

Wiley Editing Services Available to All Authors
Should you be interested, Wiley Editing Services offers expert help with manuscript, language, and format editing, along with other article preparation services. You can learn more about this service option at www.wileyauthors.com/eeo/preparation. You can also check out Wiley’s collection of free article preparation resources for general guidance about writing and preparing your manuscript at www.wileyauthors.com/eeo/prepresources.

This journal offers a number of license options for published papers; information about this is available here: https://authorservices.wiley.com/author-resources/Journal-Authors/licensing/index.html. The submitting author has confirmed that all co-authors have the necessary rights to grant in the submission, including in light of each co-author’s funder policies. If any author’s funder has a policy that restricts which kinds of license they can sign, for example if the funder is a member of Coalition S, please make sure the submitting author is aware.

Thank you for submitting your research to Molecular Ecology Resources.

Best Wishes,

Benjamin Sibbett PhD
Executive Editor, Molecular Ecology Resources
E-mail: molecol@wiley.com

### 2024-12-03 body-text-of-paper.tex
Revisions for pub/accept

%%%%%%%%%%%
\begin{figure}
\newcommand{\gspcapone}{\footnotesize
A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
individuals 1--4.  The  triangles above each of those founders denotes, in color
and text, the population of origin of each of the two genomes within each founder.
Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
It represents samples that are taken from non-founder node~7, which, from the structure
of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
on edges above and leading into each non-founder node indicate the number of
gametes that get segregated from the parent node (at the top of the edge) to the
daughter node (at the bottom of the edge).
This emphasizes that,
in dropping genes through a GSP, in contrast
to a traditional pedigree, each parent
node may segregate multiple gametes to a daughter node.  See the text for
further explanation.
Each non-founder node $x$ will have exactly two edges coming into it from above
(from its two parents), and we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
A founder individual may have, at most, 2 edges coming downward out of it, while a non-founder individual may have more or less than two edges coming downward
out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
non-founder daughter nodes, we denote the number of gametes segregated
through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example, at
non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
Finally, the purple number adjacent to the edge leading into the
sample node indicates how many sample individuals are drawn from individual
node 7. Any non-founder individual node can have a maximum of one edge leading
to a sample node.  We denote the number of samples emanating from node $x$ as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
created from a single simulation on this GSP.
}
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/F2.pdf}
\end{center}
\caption[\gspcapone]{\sometimes{\gspcapone}}
\label{fig:gsp1}
\end{figure}
%%%%%%%%%%%%
%%%%%
\begin{figure}
\newcommand{\gspcaptwo}{\footnotesize
An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $\BC_1$ in sample node~s9; and two $\BC_2$'s in sample node~s10.
}
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/AllThem.pdf}
\end{center}
\caption[\gspcaptwo]{\sometimes{\gspcaptwo}}
\label{fig:gsp2}
\end{figure}
%%%%

@@

Figure~\ref{fig:gsp1} shows an example GSP for the simple case
of simulating $F_2$ individuals. The figure and its caption
should be studied, as the following section
uses terminology that is defined in the figure caption.
It is important to understand that, apart from the founder nodes, the individual
nodes in a GSP do not necessarily represent only a single individual.  Specifically,
since we are preserving all genetic material, multiple gametes may get segregated
through each non-founder node in a GSP\@.

@@

In the \gscramble{} package there is a convenience function called {\footnotesize\tt create\_GSP()} that
will create GSPs for samples up to and including second-generation backcrosses (an example appears in Figure~\ref{fig:gsp2});
however, \gscramble{} is designed
to handle user-specified GSPs of arbitrary complexity.  When creating a unique GSP,
there are several requirements to ensure that the GSP is valid
and the simulation can execute successfully.
Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
nodes that are not parents of other non-founders, but are adjacent to a sample node, $\nonfounde$ for all
non-founder nodes that have non-founder node daughters, but no sample nodes, and $\nonfoundse$ for all
non-founder nodes that are parents of non-founder nodes and also adjacent to sample-nodes, the necessary conditions
for a valid GSP, in addition to having no inbreeding loops, are:
\begin{equation}
\begin{aligned}
\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders| &  \\
\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\
 S_x/2  &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfounds \\
 S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\
\end{aligned}
\end{equation}
In words, the separate lines above specify that 1) the number of samples
must equal the number of founders; 2) exactly two gametes must be
segregated from each founder; 3) the number of gametes to a non-founder along the 
edges from its parents must be equal; 
4) the number of gametes going downward, out of a non-founder with no sample
daughters, must equal the number of gametes coming into that non-founder;
5)  the total number of gametes entering a non-founder with only sample daughters
must be half the number of sample daughters; and 6) the number of gametes
entering non-founders with both sample daughters and non-founder daughters must
equal half the number of its sample daughters plus the sum of the number of gametes segregated
along each edge to all of its non-founder daughters.

@@

In the original paper, \citet{rizza2023limited} identified 15 $F_1$s,
1 $F_2$,
6 cutthroat backcrosses ($BC_1$--CCTs) and 
6 second-generation cutthroat backcrosses ($BC_2$--CCTs),
and 1 fish of an unknown hybrid category that may have been the
product of a $BC_2$ crossed with a $BC_1$.   In our simulations, we sought to simulate
a similar collection of hybrid individuals, so, for each replicate, we used
\gscramble{} to simulate 16 $F_1$s, 4 $F_2$s, 6 $BC_1$s, 4 $BC_2$s, 4 $BC_3$s,  4 individuals
that were the products of a $BC_1$ mating with a $BC_2$, and finally, 4 individuals that were simulated pure cutthroat. Individuals not used as founders
in \gscramble{} were retained as pure cutthroat and pure steelhead.  The goal of the simulation
was to assess how well these different hybrid categories can be distinguished from one another
when using either all 65 markers, or just the 30 markers occurring on different chromosome arms
used by  \citet{rizza2023limited}.

@@

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories. 
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair  (see
 Table~\protect\ref{tab:newhybcats} for the meanings of the names).  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/newhybs-texed-crop.pdf}
\end{center}
\caption[\nhbpcap]{\sometimes{\nhbpcap}}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to STRUCTURE or ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT (cutthroat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage (e.g., as 
implemented in {\sc NewHybrids}) 
can yield a poor estimate of the uncertainty around different inferences.

@@

Here, we present \gscramble{}, an R package implementing a simulation
approach for creating biologically informed individual 
genotypes of different hybrid classes. Two integral features of \gscramble{} are: 
1) alleles are sampled from populations without replacement and 
2) alleles are segregated based on species-specific recombination rates. 
Our results show that both 
of these features are especially important for accurately reflecting empirical data from large genetic 
marker sets (thousands of markers), 
which are common from both custom SNP-chip genotyping and next generation sequencing methods. Sampling alleles without 
replacement, and thereby preserving the genetic differences (or lack thereof) of populations is 
essential for avoiding RISPI---an issue that is exacerbated as the number of 
genetic markers increases. Additionally, we demonstrate the importance of accounting for 
the physical linkage of markers when large marker sets are used, and that treating physically linked loci as 
independent erroneously reduces the variation in the admixture 
fractions of simulated individuals.

@@

>>>>

@@

Figure~\ref{fig:gsp1} shows an example GSP for the simple case
of simulating $F_2$ individuals.
%%%%%%%%%%%
\begin{figure}
\newcommand{\gspcapone}{\footnotesize
A genomic simulation pedigree (GSP) for simulating $F_2$ individuals. See text
for detailed explanation.
}
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/F2.pdf}
\end{center}
\caption[\gspcapone]{\sometimes{\gspcapone}}
\label{fig:gsp1}
\end{figure}
%%%%%%%%%%%%
In this figure, squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
individuals 1--4.  The  triangles above each of those founders denotes, in color
and text, the population of origin of each of the two genomes within each founder.
Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
It represents samples that are taken from non-founder node~7, which, from the structure
of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
on edges above and leading into each non-founder node indicate the number of
gametes that get segregated from the parent node (at the top of the edge) to the
daughter node (at the bottom of the edge).
This emphasizes that,
in dropping genes through a GSP, in contrast
to a traditional pedigree, each parent
node may segregate multiple gametes to a daughter node.  See the text for
further explanation.
Each non-founder node $x$ will have exactly two edges coming into it from above
(from its two parents), and we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
A founder individual may have, at most, 2 edges coming downward out of it, while a non-founder individual may have more or less than two edges coming downward
out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
non-founder daughter nodes, we denote the number of gametes segregated
through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example, at
non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
Finally, the purple number adjacent to the edge leading into the
sample node indicates how many sample individuals are drawn from individual
node 7. Any non-founder individual node can have a maximum of one edge leading
to a sample node.  We denote the number of samples emanating from node $x$ as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
created from a single simulation on this GSP.
It is important to understand that, apart from the founder nodes, the individual
nodes in a GSP do not necessarily represent only a single individual.  Specifically,
since we are preserving all genetic material, multiple gametes may get segregated
through each non-founder node in a GSP\@.

@@

In the \gscramble{} package there is a convenience function called {\footnotesize\tt create\_GSP()} that
will create GSPs for samples up to and including second-generation backcrosses (an example appears in Figure~\ref{fig:gsp2});
%%%%%
\begin{figure}
\newcommand{\gspcaptwo}{\footnotesize
An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $\BC_1$ in sample node~s9; and two $\BC_2$'s in sample node~s10.
}
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/AllThem.pdf}
\end{center}
\caption[\gspcaptwo]{\sometimes{\gspcaptwo}}
\label{fig:gsp2}
\end{figure}
%%%%
however, \gscramble{} is designed
to handle user-specified GSPs of arbitrary complexity.  When creating a unique GSP,
there are several requirements to ensure that the GSP is valid
and the simulation can execute successfully.
Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
nodes that are not parents of other non-founders, but are adjacent to a sample node, $\nonfounde$ for all
non-founder nodes that have non-founder node daughters, but no sample nodes, and $\nonfoundse$ for all
non-founder nodes that are parents of non-founder nodes and also adjacent to sample-nodes, the necessary conditions
for a valid GSP, in addition to having no inbreeding loops, are:
\begin{equation}
\begin{aligned}
\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders| &  \\
\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\
 S_x/2  &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfounds \\
 S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\
\end{aligned}
\end{equation}
In words, the separate lines above specify that 1) the number of samples
must equal the number of founders; 2) exactly two gametes must be
segregated from each founder; 3) the number of gametes to a non-founder along the 
edges from its parents must be equal; 
4) the number of gametes going downward, out of a non-founder with no sample
daughters, must equal the number of gametes coming into that non-founder;
5)  the total number of gametes entering a non-founder with only sample daughters
must be half the number of sample daughters; and 6) the number of gametes
entering non-founders with both sample daughters and non-founder daughters must
equal half the number of its sample daughters plus the sum of the number of gametes segregated
along each edge to all of its non-founder daughters.

@@

In the original paper, \citet{rizza2023limited} used 30 of the 65 markers that
occur on different chromosome arms to identify 15 $F_1$s,
1 $F_2$,
6 cutthroat backcrosses ($BC_1$--CCTs) and 
6 second-generation cutthroat backcrosses ($BC_2$--CCTs),
and 1 fish of an unknown hybrid category that may have been the
product of a $BC_2$ crossed with a $BC_1$.   In our simulations, we sought to simulate
a similar collection of hybrid individuals, so, for each replicate, we used
\gscramble{} to simulate 16 $F_1$s, 4 $F_2$s, 6 $BC_1$s, 4 $BC_2$s, 4 $BC_3$s,  4 individuals
that were the products of a $BC_1$ mating with a $BC_2$, and finally, 4 individuals that were simulated pure cutthroat. Individuals not used as founders
in \gscramble{} were retained as pure cutthroat and pure steelhead.  The goal of the simulation
was to assess how well these different hybrid categories can be distinguished from one another
when using either all 65 markers, or just the 30 markers occurring on different chromosome arms
used by  \citet{rizza2023limited}.

@@

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories. 
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair  (see
 Table~\protect\ref{tab:newhybcats} for the meanings of the names).  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/newhybs-texed-crop.pdf}
\end{center}
\caption[\nhbpcap]{\sometimes{\nhbpcap}}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to STRUCTURE or ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT (cutthroat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage (e.g., as 
implemented in {\sc NewHybrids}) 
can yield a poor estimate of the uncertainty around different inferences, and this analysis
supports the choice of \citet{rizza2023limited} to limit markers to those

@@

Here, we present \gscramble{}, an R package implementing a simulation
approach for creating biologically informed individual 
genotypes of different hybrid classes. Two integral features of \gscramble{} are: 
1) alleles are sampled from populations without replacement and 
2) alleles are segregated based on species-specific recombination rates. 
Our results show that both 
of these features are especially important for accurately reflecting empirical data from large genetic 
marker sets (thousands of markers), 
which are common from both custom SNP-chip genotyping and next generation sequencing methods. Sampling alleles without 
replacement, and thereby preserving the genetic differences (or lack thereof) of populations is 
essential for avoiding RISPI---an issue that is exacerbated as the number of 
genetic markers increases. Additionally, we demonstrate the importance of accounting for 
the physical linkage of markers when large marker sets are used, and that treating physically linked loci as 
independent erroneously reduces the variation in the admixture 
fractions of simulated individuals.  In connection with this, we also stress that simply
pruning markers to eliminate or reduce linkage disequilibrium (LD)
will not solve the problem posed by physical linkage.
Genomic material is inherited in chunks that are typically orders of magnitude longer than
the typical range over which LD decays.  Thus, pairs of markers that are
on the same chromosome, are much more likely to be co-inherited than markers that
are not on the same chromosome, even if the two markers are far enough apart on the
same chromosome that they are not in LD.  That said, for some downstream analyses (after
running \gscramble{}),  LD pruning can be advantageous; for example, ADMIXTURE will run faster on a smaller, pruned data set.

@@

As noted in the methods, the GSP used to simulate with \gscramble{} must not contain
inbreeding loops, since the occurrence of such loops means genes would be
sampled with replacement from the founders, potentially inducing RISPI and incorrectly
inflating the
perceived power for admixture fraction estimation.  One might reasonably ask whether this
limits the application of \gscramble{} in small populations of conservation concern where 
inbreeding occurs; we submit that it does not.  If genetic structure between two
locales is driven by family structure (as it often may be in very small and inbred populations), that
family structure will be reflected in samples taken from each locale---the samples within a locale 
may be close relatives of one another.  If related individuals are used as founders in a GSP, then
the offspring they produce will essentially be inbred, even though the GSP itself contains no
inbreeding loops, because the founders are related.  The inclusion of related founders in this way, then, also appropriately reflects the genetic drift that is causing allele frequency differences between the locales.

### 2024-12-03 data-accessibility.tex
Update author contributions
Tiny tweaks before submitting
Revisions for pub/accept

@@

The gscramble simulation approach grew from discussions between TJS and ECA\@.
ECA, RMG, MGD, and TJS wrote the R package, `gscramble.'   Simulations of bias due to RISPI and
of variance in admixture fractions due to physical linkage were conducted by ECA.
The empirical analyses using `gscramble' were conducted
by MGD and RMG with data from TJS and by ECA with previously published data.  All authors contributed to writing and
editing the manuscript.

\section*{Data Availability Statement}

\begin{itemize}
\item Stable version of \gscramble{} available on CRAN: \url{https://CRAN.R-project.org/package=gscramble}
\item Development version and entire revision history on GitHub: \url{https://github.com/eriqande/gscramble}
\item Online version of all package documentation: \url{https://eriqande.github.io/gscramble}
\item Data and code repository with simulation analyses and steelhead-cutthroat analyses: \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} (To be transferred to Dryad or Zenodo upon acceptance).
\item Data and code repository for pig analyses: \url{https://github.com/mgdesaix/gscramble-paper-pigs} (To be transferred to Dryad or Zenodo upon acceptance)
\end{itemize}
\mbox{}

>>>>

\section*{Data Accessibility and Benefit-Sharing}

\subsection*{Data Accessibility Statement}

A stable version of the `gscramble' R package is available on CRAN: \url{https://CRAN.R-project.org/package=gscramble}. The development version and entire revision history of
\gscramble{} is available on GitHub: \url{https://github.com/eriqande/gscramble}, and all package documentation is available online in HTML format: \url{https://eriqande.github.io/gscramble}

Data and code for the simulation analyses and steelhead-cutthroat analyses are available from \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} and are archived at \url{https://doi.org/10.5281/zenodo.14270660}.  Data and code for the pig analyses are available from \url{https://github.com/mgdesaix/gscramble-paper-pigs} and are archived
at \url{Get-from-Matt}.

\subsection*{Benefit-sharing Statement}

Benefits Generated: Benefits from this research accrue from the sharing of our data and results on public databases and the distribution of a novel and useful R package as described above.

@@

The gscramble simulation approach grew from discussions between TJS and ECA\@.
ECA, RMG, MGD, and TJS wrote the R package, `gscramble.'   Simulations of bias due to RISPI and
of variance in admixture fractions due to physical linkage were conducted by ECA.
The empirical analyses using `gscramble' were conducted by MGD, RMG, TJS, and ECA with previously published data.
All authors contributed to writing and editing the manuscript.
